---
title: "Lab 3"
author: "Shawn Santo"
date: "09/16/19"
output: 
  html_document:
    css: lab3.css
    toc: true
    number_sections: false
    toc_float: true
    df_print: paged
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      comment = NA,
                      fig.width=8, 
                      fig.height=6)
```

## Introduction

Today we will investigate a data set from 
[http://www.gapminder.org](http://www.gapminder.org), 
a site which contains a wealth of data and visualizations related to 
health, wealth, population, etc. of countries of the world.

The data is available in package `gapminder` as a tibble object named
`gapminder`. Install the package in your console with 
`install.packages("gapminder")` and load it along with other packages by

```{r packages}
library(gapminder)
library(tidyverse)
library(RColorBrewer)
```

```{r view-data}
gapminder
class(gapminder)
```

We'll create an object named `gapminder_df` that only has class `data.frame`.
Hence, we remove class `tbl_df` and `tbl`.

```{r class-change}
gapminder_df <- gapminder
attr(gapminder_df, which = "class") <- "data.frame"
class(gapminder_df)
```


Object `gapminder` has six variables:

- **country**: country of the world, factor
- **continent**: continent of the world, factor
- **year**: year, integer
- **lifeExp**: median life expectancy, double
- **pop**: population, integer
- **gdpPercap**: GDP per capita, double

To see the HTML file for Lab 3 
visit http://rpubs.com/shawnsanto/sta523-fa19-lab3

## Exercises

### Data frame manipulations

For exercises 1 - 5 perform the task using both `gapminder` and `gapminder_df`.
Focus on the differences in results between the two data 
frames to get a better understanding of the tibble object. 

1. Extract the third row.

2. Extract the third column.

3. Extract the third column so the result is a data frame.

4. Extract the first and last columns.

5. Extract the last 20 rows without using the fact that you know there are
   1704 rows of data.

6. What years are in the data set? *Hint:* `unique`.

7. How many countries since 1960 had a median life expectancy of at least 80?

8. Create a new variable named `pop_scale` for `gapminder_df` that is defined
   as `pop / 10000`.
   
9. What was the mean life expectancy for Europe in 1957? How about in 2007?

10. Which country had the highest median life expectancy in 1957? How about in
    2007?
    
### Plots {.tabset .tabset-fade .tabset-pills}

Use object `gapminder` to recreate the plots using functions in package
`ggplot2`. To see the plots you will need to 
visit http://rpubs.com/shawnsanto/sta523-fa19-lab3

#### Plot 1 { - .tabset .tabset-fade .tabset-pills}

##### Plot {-}

```{r}


```

##### Hints {-}

Non-obvious plot features

- `color = "blue", alpha = .2`
- `base_size = 20`

#### Plot 2 { - .tabset .tabset-fade .tabset-pills}

##### Plot {-}

```{r fig.height=12}



```

##### Hints {-}

Non-obvious plot features

- Filter `gapminder` for `Americas` and years 1957, 1982, 2007.
- `show.legend = FALSE`

#### Plot 3 { - .tabset .tabset-fade .tabset-pills}

##### Plot {-}

```{r fig.height=12, fig.width=10}



```

##### Hints {-}

Non-obvious plot features

- Filter `gapminder` so it does not contain `Oceania`
- `colors = brewer.pal(9, "Reds")`

#### Plot 4 { - .tabset .tabset-fade .tabset-pills}

##### Plot {-}

```{r echo=FALSE}



```

##### Hints {-}

Non-obvious plot features

- Filter `gapminder` so it does not contain `Oceania`
- `alpha = .3`
- Use a linear model fit on the entire data
- Transfer `x` to a `log10` scale